Comparison of Terrain Following and Cut Cell Grids with a Non-Hydrostatic Model

### 2015-01-29 msc-publication.tex
add more references in AMS bibtex format

\documentclass{ametsoc}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{siunitx}

@@

Builds on the work of \citet{schaer2002} (MWR), \textcite{klemp2011} (MWR), \textcite{weller-shahrokhi2014} (MWR)

Results also compared with \textcite{melvin2010} (QRJ), maybe \textcite{zaengl2012} (MWR), \textcite{good2014} (Atmos Sci Letters)

@@

\section{Results}
\begin{itemize}
	\item \TODO{How much do I need to say about the model?  Just cite \textcite{weller-shahrokhi2014}?}
	\item \TODO{Include BTF equations from \textcite{galchen-somerville1975} and SLEVE equations from \textcite{schaer2002}?}
	\item \TODO{Explain how SnapCol grid is constructed?}
\end{itemize}

@@

Analysis:
\begin{itemize}
	\item Upwind-biased cubic is qualitatively better than Sch\"ar's 4th order leapfrog on BTF grid (figure~\ref{fig:advection})
	\item SnapCol error, min and max is almost identical to noOrography (unsurprisingly) (table~\ref{tab:advection})  This agrees with \textcite{good2014}
	\item SLEVE error, min and max are part-way between BTF and noOrography (again, unsurprisingly)
	\item noOrography, BTF and SLEVE min and max comparable to \textcite{schaer2002}
	\item Conclusion: advection more accurate on more regular grids, but nevertheless satisfactory on all grids
\end{itemize}

@@

Analysis:
\begin{itemize}
	\item Spurious $w$ significantly less than those found by \textcite{klemp2011} for BTF: \SI{0.35}{\meter\per\second} vs \SI{10}{\meter\per\second}
	\item SnapCol $w$ (\SI{1e-3}{\meter\per\second}) significantly lower and energy conservation better than BTF or SLEVE
	\item Formulation is not energy conserving
	\item \TODO{Could also compare with \textcite{zaengl2012}, \textcite{good2014}?}
	\item Conclusion: non-orthogonality is a significant cause of numerical error in this test but, again, satisfactory on all grids
\end{itemize}

\begin{figure}
	\includegraphics[height=3in]{resting-atmosphere-w.png}
%
	\caption{Comparing our results with \textcite{klemp2011}'s}
	\label{fig:resting}
\end{figure}

@@

Analysis:
\begin{itemize}
	\item $w$ contours visually similar on all grids, agree with \textcite{melvin2010} (Figure~\ref{fig:gw-w})
	\item \TODO{don't know if divergence is relevant -- it's greater on the SnapCol grid (MSc dissertation Figure 4.15e)}
	\item $\theta$ anomalies similar on all grids EXCEPT...
	\item ... on SnapCol grid in lee of mountain near the ground. (Figure~\ref{fig:gw-theta})
	\item \TODO{worth mentioning implications of this?  1. reduced stability, although not enough to create vertical motion in this instance 2. model has no viscosity so thermal mixing should not occur in theory}
	\item Lorenz computation mode has been excited because Exner sample line at $x = \SI{50}{\kilo\meter}$ is in hydrostatic balance (Figure~\ref{fig:gw-exner-theta})
	\item \TODO{We can speculate on what excites the computational mode, but perhaps better not to?}
	\item Little/no evidence of small cell problem -- \TODO{don't know that we can say much here without another vertical momentum test; our hypothesis about the quasi-horizontal flow isn't backed up with a test case}
	\item Conclusion: results similar on all grids, agree with literature, except for Lorenz computational mode on SnapCol grid
\end{itemize}

@@

\begin{figure}
	\includegraphics[height=3in]{gw-theta.png}
	\includegraphics[height=3in]{gw-theta-zoom.png}
%
	\caption{Gravity wave thermal anomalies from initial thermal profile}
	\label{fig:gw-theta}
\end{figure}

\begin{figure}
	\includegraphics[height=3in]{gw-exner-theta.png}
%
	\caption{Vertical sample line of Exner function of pressure and theta to demonstrate Lorenz computational mode}
	\label{fig:gw-exner-theta}
\end{figure}

\section{Conclusions}
\begin{enumerate}
	\item BTF grid isn't as bad as people say it is.  We found that:
	\begin{itemize}
		\item Upwind-biased cubic advection scheme accurately advects a tracer in non-divergent flows (Figure~\ref{fig:advection}, Table~\ref{tab:advection})
		\item spurious vertical velocities are small in resting atmosphere (Figure~\ref{fig:resting})
		\item gravity waves results visually as good as reference solution from \textcite{melvin2010} (figure~\ref{fig:gw-w})
	\end{itemize}

@@

\begin{table}[t]
\caption{Min, max and error norms compared with \textcite{schaer2002}}
\label{tab:advection}
%
\begin{center}
\begin{tabular}{ l l l l l l }
\hline\hline
& \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{2}{c}{Sch\"ar 4th order} \\
& $\ell^2$ error & min & max & min & max \\
\hline
Analytic  & 0 & 0 & 1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
BTF 	  & TODO & TODO & TODO & \num{-0.058} & \num{1.001} \\
SnapCol   & TODO & TODO & TODO & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
noOrography & TODO & TODO & TODO & \num{-0.002} & \num{0.984} \\
\hline
\end{tabular}
\end{center}
\end{table}

>>>>

\documentclass[twocol]{ametsoc}
\journal{mwr}

\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{siunitx}

@@

Builds on the work of \citet{schaer2002}, \citet{klemp2011}, \citet{weller-shahrokhi2014} (all MWR)

Results also compared with \citet{melvin2010} (QRJ), maybe \textcite{zaengl2012} (MWR), \citet{good2014} (Atmos Sci Letters)

@@

\section{Results}
\begin{itemize}
	\item \TODO{How much do I need to say about the model?  Just cite \citet{weller-shahrokhi2014}?}
	\item \TODO{Include BTF equations from \textcite{galchen-somerville1975} and SLEVE equations from \citet{schaer2002}?}
	\item \TODO{Explain how SnapCol grid is constructed?}
\end{itemize}

@@

Analysis:
\begin{itemize}
	\item Upwind-biased cubic is qualitatively better than Sch\"ar's 4th order leapfrog on BTF grid (figure~\ref{fig:advection})
	\item SnapCol error, min and max is almost identical to noOrography (unsurprisingly) (table~\ref{tab:advection})  This agrees with \citet{good2014}
	\item SLEVE error, min and max are part-way between BTF and noOrography (again, unsurprisingly)
	\item noOrography, BTF and SLEVE min and max comparable to \citet{schaer2002}
	\item Conclusion: advection more accurate on more regular grids, but nevertheless satisfactory on all grids
\end{itemize}

@@

Analysis:
\begin{itemize}
	\item Spurious $w$ significantly less than those found by \citet{klemp2011} for BTF: \SI{0.35}{\meter\per\second} vs \SI{10}{\meter\per\second}
	\item SnapCol $w$ (\SI{1e-3}{\meter\per\second}) significantly lower and energy conservation better than BTF or SLEVE
	\item Formulation is not energy conserving
	\item \TODO{Could also compare with \textcite{zaengl2012}, \citet{good2014}?}
	\item Conclusion: non-orthogonality is a significant cause of numerical error in this test but, again, satisfactory on all grids
\end{itemize}

\begin{figure}
	\includegraphics[height=2in]{resting-atmosphere-w.png}
%
	\caption{Comparing our results with \citet{klemp2011}'s}
	\label{fig:resting}
\end{figure}

@@

Analysis:
\begin{itemize}
	\item $w$ contours visually similar on all grids, agree with \citet{melvin2010} (Figure~\ref{fig:gw-w})
	\item \TODO{don't know if divergence is relevant -- it's greater on the SnapCol grid (MSc dissertation Figure 4.15e)}
	\item $\theta$ anomalies similar on all grids EXCEPT...
	\item ... on SnapCol grid in lee of mountain near the ground. (Figure~\ref{fig:gw-theta})
	\item \TODO{worth mentioning implications of this?  1. reduced stability, although not enough to create vertical motion in this instance 2. model has no viscosity so thermal mixing should not occur in theory}
	\item Lorenz computation mode has been excited because Exner sample line at $x = \SI{50}{\kilo\meter}$ is in hydrostatic balance (Figure~\ref{fig:gw-exner-theta})
	\item \TODO{We can speculate on what excites the computational mode, but perhaps better not to?}
	\item Little/no evidence of small cell problem -- \TODO{don't know that we can say much here without another vertical momentum test; our hypothesis about the quasi-horizontal flow isn't backed up with a test case}
	\item Conclusion: results similar on all grids, agree with literature, except for Lorenz computational mode on SnapCol grid
\end{itemize}

@@

\begin{figure}
	\includegraphics[height=3in]{gw-theta.png}
	\includegraphics[height=2in]{gw-theta-zoom.png}
%
	\caption{Gravity wave thermal anomalies from initial thermal profile}
	\label{fig:gw-theta}
\end{figure}

\begin{figure}
	\includegraphics[height=2in]{gw-exner-theta.png}
%
	\caption{Vertical sample line of Exner function of pressure and theta to demonstrate Lorenz computational mode}
	\label{fig:gw-exner-theta}
\end{figure}

\section{Conclusions}
\begin{enumerate}
	\item BTF grid isn't as bad as people say it is.  We found that:
	\begin{itemize}
		\item Upwind-biased cubic advection scheme accurately advects a tracer in non-divergent flows (Figure~\ref{fig:advection}, Table~\ref{tab:advection})
		\item spurious vertical velocities are small in resting atmosphere (Figure~\ref{fig:resting})
		\item gravity waves results visually as good as reference solution from \citet{melvin2010} (figure~\ref{fig:gw-w})
	\end{itemize}

@@

\begin{table}[t]
\caption{Min, max and error norms compared with \citet{schaer2002}}
\label{tab:advection}
%
\begin{center}
\begin{tabular}{ l l l l l l }
\hline\hline
& \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{2}{c}{Sch\"ar 4th order} \\
& $\ell^2$ error & min & max & min & max \\
\hline
Analytic  & 0 & 0 & 1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
BTF 	  & TODO & TODO & TODO & \num{-0.058} & \num{1.001} \\
SnapCol   & TODO & TODO & TODO & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
noOrography & TODO & TODO & TODO & \num{-0.002} & \num{0.984} \\
\hline
\end{tabular}
\end{center}
\end{table}
